Chord: A Versatile Platform for Program Analysis

### 2011-03-15 main/doc/datalog_analysis.tex
more documentation

A common way to rapidly prototype program analyses in Chord is using a declarative
logic-programming language called Datalog.  This chapter describes everything about
Datalog analysis in Chord.
Section \ref{sec:writing-datalog-analysis} explains how to write a Datalog analysis,
Section \ref{sec:running-datalog-analysis} explains how to run it, and
Section \ref{sec:tuning-datalog-analysis} explains how to tune its performance.
Finally, Section \ref{sec:bdd} explains the representation of BDDs (Binary Decision Diagrams)
which are a data structure central to executing Datalog analyses.

@@

A Datalog analysis declares a bunch of input/output program relations,
each over one or more program domains, and provides a bunch of rules (constraints)
specifying how to compute the output relations from the input relations.
An example of such an analysis is as follows:

\begin{verbatim}
    # name=datarace-dlog

@@

    # Analysis constraints
    hasWrite(e1,e2) :- write(e1).
    hasWrite(e1,e2) :- write(e2).
    candidate(e1,e2) :- field(e1,f), field(e2,f), hasWrite(e1,e2), e1 <= e2.
    datarace(t1,e1,t2,e2) :- candidate(e1,e2), reach(t1,e1), reach(t2,e2), \
        alias(e1,e2), escape(e1), escape(e2), unguarded(t1,e1,t2,e2).
\end{verbatim}

@@

Each relation is represented symbolically (as opposed to explicitly)
using a graph-based data structure called a Binary Decision Diagram (BDD for short).
Each domain containing N elements is assigned log2(N) BDD variables.
The size of a BDD and the efficiency of operations on it depends heavily
on the order of these BDD variables.
The {\tt .bddvarorder <...>} line in the Datalog analysis enables the Datalog
analysis writer to specify this order.
It must list all domains along with their numerical suffixes, separated
by {\tt \_} or {\tt x}.
Using a {\tt \_} between two domains, such as {\tt T0\_T1}, means that the BDD variables assigned
to domain {\tt T0} precede those assigned to domain {\tt T1} in the BDD variable
order for this Datalog analysis.
Using a {\tt x} between two domains, such as {\tt E0xE1}, means that the
BDD variables assigned to domains {\tt E0} and {\tt E1}
will be interleaved in the BDD variable order for this Datalog analysis.
See Section \ref{chap:bdd-representation} for more details on BDD representations.

>>>>

A common way to rapidly prototype program analyses in Chord is using a declarative
logic-programming language called Datalog.  This chapter describes all aspects of
Datalog analysis in Chord.
Section \ref{sec:writing-datalog-analysis} explains how to write a Datalog analysis,
Section \ref{sec:running-datalog-analysis} explains how to run it, and
Section \ref{sec:tuning-datalog-analysis} explains how to tune its performance.
Finally, Section \ref{sec:bdd} explains the representation of BDDs (Binary Decision Diagrams)
which are a data structure central to executing Datalog analyses.

@@

A Datalog analysis declares a bunch of input/output program relations,
each over one or more program domains, and provides a bunch of rules (constraints)
specifying how to compute the output relations from the input relations.
An example of such an analysis is shown in Figure \ref{fig:datalog-analysis}.

\begin{figure}
{\small
\begin{verbatim}
    # name=datarace-dlog

@@

    # Analysis constraints
    hasWrite(e1,e2) :- write(e1).
    hasWrite(e1,e2) :- write(e2).
    candidate(e1,e2) :- field(e1,f), field(e2,f), hasWrite(e1,e2), e1 <= e2.
    datarace(t1,e1,t2,e2) :- candidate(e1,e2), reach(t1,e1), reach(t2,e2), \
        alias(e1,e2), escape(e1), escape(e2), unguarded(t1,e1,t2,e2).
\end{verbatim}
}
\label{fig:datalog-analysis}
\caption{An Example Datalog Analysis.}
\end{figure}

@@

Each relation is represented symbolically (as opposed to explicitly)
using a graph-based data structure called a Binary Decision Diagram (BDD for short).
Each domain containing N elements is assigned log2(N) BDD variables.
The size of a BDD and the efficiency of operations on it depends heavily
on the order of these BDD variables.
The {\tt .bddvarorder <...>} line in the Datalog analysis enables the Datalog
analysis writer to specify this order.
It must list all domains along with their numerical suffixes, separated
by {\tt \_} or {\tt x}.
Using a {\tt \_} between two domains, such as {\tt T0\_T1}, means that the BDD variables assigned
to domain {\tt T0} precede those assigned to domain {\tt T1} in the BDD variable
order for this Datalog analysis.
Using a {\tt x} between two domains, such as {\tt E0xE1}, means that the
BDD variables assigned to domains {\tt E0} and {\tt E1}
will be interleaved in the BDD variable order for this Datalog analysis.
See Section \ref{sec:bdd} for more details on BDD representations.
